Learning Reusable Components for Neurally--Guided Bayesian Program Learning

### 2018-05-13 nips_2018.tex
string example structure

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we provide a generic
 set of programming primitives in a Lisp-like language,
 including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(join '' (split c s))}\\
   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
   &\hspace{0.5cm} \emph{(add k to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{wrapfigure}

@@

\subsection{Symbolic Regression}\label{regressionSection}
\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
As a simple first example of how our algorithm works, we apply \system
to symbolic regression problems.  For each of these problems, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
200 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  Our recognition model here is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of continuous parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\subsection{Programs that manipulate sequences}\label{sequences}
We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
For both these domains we initially provide the system with a generic
minimal set of programming primitives,
and use a bidirectional GRU~\cite{cho2014learning} for the recognition model.
The initial set of primitives includes routines commonly found in Lisp or Scheme interpreters: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsubsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other system successful.
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
At first, \system cannot find any correct programs for most of the tasks.
It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

>>>>

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we provide a generic
 set of programming primitives in a Lisp-like language,
 including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
   &\hspace{0.5cm} \emph{(add k to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{wrapfigure}

@@

\subsection{Symbolic Regression}\label{regressionSection}
\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
As a simple first example of how our algorithm works, we apply \system
to symbolic regression problems.  For each of these problems, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
200 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  Our recognition model here is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\subsection{Programs that manipulate sequences}\label{sequences}
We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
For both these domains we initially provide the system with a generic
minimal set of programming primitives,
and use a bidirectional GRU~\cite{cho2014learning} for the recognition model.
The initial set of primitives includes routines commonly found in Lisp or Scheme interpreters: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsubsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other system successful.
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
At first, \system cannot find any correct programs for most of the tasks.
It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.
